Predicting complex diseases: performance and robustness

### 2017-10-26 paper/draft.tex
save

@@

@@

@@

[TIMING: PRS -> 0; logit -> 500-600 with small increase when more causal variants, logit-triple x3]

[AUC: decrease with M, PRS-all never good, logit-triple better for fancy and almost as good for simple]

[\% Cases: Linear with AUC, Attention perc cases is == 30%]

@@

@@

Authors acknowledge Grenoble Alpes Data Institute, supported by the French National Research Agency under the ``Investissements d?avenir'' program (ANR-15-IDEX-02) and the LabEx PERSYVAL-Lab (ANR-11-LABX-0025-01).

@@

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

[MONTRER UN MANHATTAN PLOT POUR JUSTIFIER 30 IN HLA]

@@

[ALSO TRIPLE LOGIT]

@@

First, we show that AUC is highly correlated with the percentage of cases in either the 10\% or 20\% largest PRSs (Figure \ref{fig:AUC-corr}). So, we report only predictive performance in terms of AUC in following results.

As for the comparison of our methods with the P+T method, we report AUCs greater than 95\%, which is not obtained by the upper bound of the P+T method in any scenario. Our penalized regression method consistently provides better predictive performance than the P+T method. Our method is parallelized and requires only less than 10 minutes to run (with 6 physical cores). Moreover, it results in sparse models (...).
The P+T runs in only a few seconds because we don't consider the GWAS computation as part of the execution time of the method because this is generally executed by others and then the resulting summary statistics are used to construct a PRS.

As for the comparison between our two penalized logistic regression, the classic one (called ``simple'') and the one using additional features coding for recessive and dominant effects (called ``triple''), we report that the ``triple'' method is nearly as good as the ``simple'' one when there are only linear effects and can lead to significantly greater results when there are also dominant and interactions effects. Yet, the ``triple'' solution takes approximately 3 more times to run.

As for general results, we report a decrease of AUC when the number of causal variants $M$ is increasing.

[TROUVER DES AUTRES: \% of decrease for h2 0.8 -> 0.5? Pas bon d'utliser tous les SNPs]

[REPLICATE INDIVIDUALS TO SHOW IF CAN BE USED ON LARGER DATA]

@@

In this comparative study, we proposed other methods as a replacement to the widely used P+T method. 
We show that jointly estimating the SNP effects using efficient penalized logistic regressions enable to achieve really good performance as reported in previous studies [REF GAD + AUTRE?], better than with the P+T method.
We are well aware that the P+T method is widely used because it is simple and can use summary statistics already learned from large genotype data, which is not the case for our penalized logistic regressions. 
Yet, in the case of prediction of disease to be used for clinical interventions, best possible methods should be used, even if it has to be used by large consortia only.

One does not simply identify high-risk individuals (Figure \ref{fig:meme}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Authors acknowledge Grenoble Alpes Data Institute, supported by the French National Research Agency under the ``Investissements d'avenir'' program (ANR-15-IDEX-02) and the LabEx PERSYVAL-Lab (ANR-11-LABX-0025-01). We are also grateful to Félix Balazard for useful discussions about T-Trees.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{AUC-corr}}
\caption{Percentage of cases in the 2 highest deciles of PRSs as a function of AUC.}\label{fig:AUC-corr}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h]
\centerline{\includegraphics[width=\textwidth]{meme}}
\caption{Meme to use on Twitter when promoting the preprint?}\label{fig:meme}
\end{figure}
